Corpus Experiments for Word Embeddings

### 2015-08-18 article/cooccurrence-noise-experiment.tex
separate files for article text samples, document frequency function

\texttt{\textcolor{red}{CAT\_3} the domestic \textcolor{red}{CAT\_3} was first classified as felis \textcolor{red}{CAT\_2} catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one \textcolor{red}{CAT\_2} whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated \textcolor{red}{CAT\_2} canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time \textcolor{red}{CAT\_3} at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

>>>>

\texttt{the \textcolor{red}{CAT\_2} domestic \textcolor{red}{CAT\_3} was \textcolor{red}{CAT\_3} first \textcolor{red}{CAT\_2} classified as felis catus\newline the semiferal \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is \textcolor{red}{CAT\_1} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} fancier \textcolor{red}{CAT\_1} organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only \textcolor{red}{CAT\_1} individuals of the same breed\newline the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye \textcolor{red}{CAT\_3} sockets\newline another unusual feature \textcolor{red}{CAT\_2} is that the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant \textcolor{red}{CAT\_1} over the others\newline ...\newline the domestic dog \textcolor{red}{CAT\_3} canis \textcolor{red}{CAT\_1} lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for \textcolor{red}{CAT\_3} people such as \textcolor{red}{CAT\_3} hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the \textcolor{red}{CAT\_2} time at \textcolor{red}{CAT\_1} which female dogs will have \textcolor{red}{CAT\_2} their first \textcolor{red}{CAT\_3} estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-08-18 article/main.tex
separate files for article text samples, document frequency function

For simplicity, only the CBOW method with a single set of hyperparameters is considered.
Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $200$, and $10$ passes through the corpus.
Subsampling was not used so that influence of word frequency could be more clearly discerned.
Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.

@@

\begin{subsection}{Tokens derived from existing words}
A small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
To this set of words we add the word \word{the}, in order to include a high-frequency stopword.
The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.
This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 200.  
A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{200}}$ exceeds the maximum occurrence of any token in the corpus. 
Figure \ref{fig:word-frequency-experiment-text-cat} illustrates the effect of these modifications on a sample text, with a family of tokens \word{CAT\_i}, derived from the word \word{cat}.

@@

\begin{figure}
	\input{word-frequency-experiment-text-cat.tex}
	\caption{Text modified for the word frequency experiment, where the word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
	\label{fig:word-frequency-experiment-text-cat}
\end{figure}

@@

\begin{subsection}{Tokens derived from an introduced, meaningless word}
	A high-frequency, purely meaningless word is included in the experiment for comparison.
	We choose to introduce a new, entirely meaningless token \word{VOID} into the corpus, rather than choose an existing word whose meaninglessness is only supposed.
	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
	The co-occurrence distribution of this introduced token thus coincides with the global frequency distribution.
	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
Figure \ref{fig:word-frequency-experiment-text-void} illustrates the effect of this modifications on a sample text.

\begin{figure}\label{fig:word-frequency-experiment-text-void}
	\input{word-frequency-experiment-text-void.tex}
	\caption{Text modified for the word frequency experiment, where the
	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
\end{figure}
\end{subsection}

@@

% original text after modifications for the noise experiment with n=3 lambda 5/ 6 for cat
\begin{figure}
	\input{cooccurrence-noise-experiment.tex}
	\caption{Text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
\label{fig:cooccurrence-noise-experiment-text}
\end{figure}

>>>>

For simplicity, only the CBOW method with a single set of hyperparameters is considered.
Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $200$, and $10$ passes through the corpus.
Subsampling was not used so that influence of word frequency could be more clearly discerned.
Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.
This cut-off results in a vocabulary of $62025$ words (only unigrams were considered).

@@

\begin{subsection}{Tokens derived from existing words}\label{WFVEexisting}
A small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
To this set of words we add the word \word{the}, in order to include a high-frequency stopword.
The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.
This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 200.  
A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{200}}$ exceeds the maximum occurrence of any token in the corpus. 
Figure \ref{fig:word-frequency-experiment-text-cat} illustrates the effect of these modifications on a sample text, with a family of tokens \word{CAT\_i}, derived from the word \word{cat}.

@@

\begin{figure}
	\input{word-frequency-experiment-text-cat.tex}
	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEexisting}, where the
	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
	\label{fig:word-frequency-experiment-text-cat}
\end{figure}

@@

\begin{subsection}{Tokens derived from an introduced, meaningless word}\label{WFVEmeaningless}
	A high-frequency, purely meaningless word is included in the experiment for comparison.
	We choose to introduce a new, entirely meaningless token \word{VOID} into the corpus, rather than choose an existing word whose meaninglessness is only supposed.
	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
	The co-occurrence distribution of this introduced token thus coincides with the global frequency distribution.
	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
Figure \ref{fig:word-frequency-experiment-text-void} illustrates the effect of this modifications on a sample text.

\begin{figure}\label{fig:word-frequency-experiment-text-void}
	\input{word-frequency-experiment-text-void.tex}
	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEmeaningless}, where $\lambda=0.5$ and $n=20$. For illustrative purposes, the token \word{VOID} was here interspersed with a global frequency of $0.05$.}
\end{figure}
\end{subsection}

@@

\begin{figure}
	\input{cooccurrence-noise-experiment.tex}
	\caption{A text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
\label{fig:cooccurrence-noise-experiment-text}
\end{figure}

### 2015-08-18 article/word-frequency-experiment-text-cat.tex
separate files for article text samples, document frequency function

\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_5} is one whose ancestry is recorded by a \textcolor{red}{CAT\_3} fancier organization\newline a purebred \textcolor{red}{CAT\_4} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

>>>>

\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_3} a mostly outdoor \textcolor{red}{CAT\_7} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_3} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}

### 2015-08-18 article/word-frequency-experiment-text-void.tex
separate files for article text samples, document frequency function

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

>>>>

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline \textcolor{red}{VOID\_1} dogs perform many roles for people such as hunting \textcolor{red}{VOID\_1} herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction\newline ...\newline rabbits are small mammals in the family leporidae\newline the male is called a buck and the female is a doe\newline the european rabbit has been introduced to many places around the world\newline more than half \textcolor{red}{VOID\_1} the worlds rabbit population resides in north america\newline rabbits are \textcolor{red}{VOID\_1} prey animals and are therefore constantly aware of their surroundings\newline rabbits are most active at dawn and dusk\newline as with other prey animals rabbits often sleep with their eyes open}
